---
title: "os"
author: "cohort selection"
date: "2024-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
vars <- c( "gender", "age_group", "combined_ecog", "issstage", "asct", "risk_group",
          "sesindex2015_2019","year_diag_group", "hemoglobin","marker_calcium_result",
          "marker_creatinie_result","practicetype",
         "Deletion.13q", "Ploidy")
os_data <- read.csv("os_data.csv")
os_data <-  os_data %>% mutate(combined_race_ethnicity = factor(combined_race_ethnicity,
                                            levels = c("White", 
                                                       "African American",
                                                       "Asian",
                                                       "Hispanic or Latino",
                                                       "Other Races",
                                                       "Unknown/Missing")))
os_data <- os_data %>% select("combined_race_ethnicity", "gender", "age_group", "combined_ecog", "issstage", "asct", "risk_group",
          "sesindex2015_2019","year_diag_group", "hemoglobin","marker_calcium_result",
          "marker_creatinie_result","practicetype",
         "Deletion.13q", "Ploidy","OS","event")
# Assume new_dat is your data frame
```

### Asian
```{r}
#install.packages("mice")
library(mice)
# set "Unknown/missing" as NA
os_data_2 <- os_data # os_data_2 is a back up dataset
os_data_2$combined_race_ethnicity[os_data_2$combined_race_ethnicity == "Unknown/Missing"] <- NA
# set anemia, renal deficiency, .. as factors
os_data_2$hemoglobin <- as.factor(os_data_2$hemoglobin)
os_data_2$marker_calcium_result <- as.factor(os_data_2$marker_calcium_result)
os_data_2$marker_creatinie_result <- as.factor(os_data_2$marker_creatinie_result)

# only include covariates of interest
os_data_2 <- os_data_2 %>% select("combined_race_ethnicity", 
                                  "gender", "age_group", "combined_ecog", "issstage", "asct", "risk_group",
                                  "sesindex2015_2019","year_diag_group", "hemoglobin","marker_calcium_result",
                                  "marker_creatinie_result","practicetype", "Deletion.13q", "Ploidy","OS","event")

# # data imputation
# 22 imputated datasets, random seed set into "123"
imputated_data <- mice(os_data_2,method =c("polyreg"), m =22, seed =123)
imputed_list <- complete(imputated_data, action = "all",include =TRUE)
races_to_keep <- c("White", "Asian" )
filtered_imputed_list <- lapply(imputed_list, function(df){
  df_filtered <- df[df$combined_race_ethnicity %in% races_to_keep,]
  df_filtered <- df_filtered %>% mutate(treatment = ifelse(combined_race_ethnicity == "White",0,1))
  return(df_filtered)
})

# To view
filtered_imputed_list[[22]]
sum(filtered_imputed_list[[22]]$combined_race_ethnicity == "Asian")
sum(filtered_imputed_list[[22]]$combined_race_ethnicity == "White")

```

```{r}
missing_idx <- which(is.na(os_data_2$combined_race_ethnicity))
length(missing_idx)  # Total number of NAs to check against

# Extract imputed race/ethnicity values at the missing positions across 22 datasets
imputed_values <- sapply(imputed_list, function(df) df$combined_race_ethnicity[missing_idx])

# Convert to long format (vector of all imputed values)
all_imputed_values <- as.vector(imputed_values)

# Tabulate counts of each category
imputed_counts <- table(all_imputed_values)

# Average across imputations
avg_counts <- round(imputed_counts / 22, 2)

# Proportions
total_imputed <- length(missing_idx)
avg_props <- round(avg_counts / total_imputed, 4)

# Combine into a data frame
result <- data.frame(
  Race_Ethnicity = names(avg_counts),
  Avg_Count_Assigned = as.numeric(avg_counts),
  Avg_Proportion = as.numeric(avg_props)
)

print(result)
```

```{r}
library(tableone)
# Calculate SMD for the unmatched data
for (i in 1:22) { 
    table1 <- CreateTableOne(vars = vars, strata = "treatment", data =  filtered_imputed_list[[i]], test = FALSE, smd = TRUE)

    smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
  imputation_results[[i]] <- smd_values
}
all_smds <- do.call(rbind, imputation_results)

all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)

# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()


# average SMDs across imputations for the unmactched data 
tabUnmatched <- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))
print(tabUnmatched)

```


# Crude 
```{r}
library(broom)
# Assuming you have multiple imputed datasets in a list called `imputed_list` 
num_imputations <- length(filtered_imputed_list) 
# Create placeholders to store coefficients and standard errors 
coef_list <- list() 
se_list <- list() 
# Fit Cox model on each imputed dataset 
for (i in 1:num_imputations) { 
  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data = filtered_imputed_list[[i]]) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
} 

# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
crude_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(crude_results)

```
# Covariate Adjustment
```{r}
# Assuming you have multiple imputed datasets in a list called `imputed_list` 
num_imputations <- length(filtered_imputed_list) 
# Create placeholders to store coefficients and standard errors 
coef_list <- list() 
se_list <- list() 
# Fit Cox model on each imputed dataset 
for (i in 1:num_imputations) { 
  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment) + age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy, data = filtered_imputed_list[[i]]) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
  } 
# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
cov_adj_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI , se =pooled_se , loghr =mean_coefs )
print(cov_adj_results)
```
```{r}
## Construct a data frame containing variable name and SMD from all methods
dataPlot <- data.frame(variable   = tabUnmatched$Variable,
                      Unmatched  = as.numeric(tabUnmatched$Mean_SMD)
                       )


## Create long-format data for ggplot2
dataPlotMelt <- melt(data          = dataPlot,
                     id.vars       = c("variable"),
                     variable.name = "Method",
                     value.name    = "SMD")

library(forcats)

dataPlotMelt$Method <- fct_recode(dataPlotMelt$Method,
  "Original Data" = "Unmatched"
)

## Order variable names by magnitude of SMD
varNames <- as.character(dataPlot$variable)[order(dataPlot$Unmatched,decreasing = FALSE)]

## Order factor levels in the same order
dataPlotMelt$variable <- factor(dataPlotMelt$variable,
                                levels = varNames)
```


# Nerest Neighbor Matching - With Caliper = 0.2

```{r}
imputation_results <- list()
coef_list <- list()
se_list <- list()
matching_summary <- list() 
for (i in 1:22) { 
  # Access the i-th imputed dataset 
 imputed_subset <- filtered_imputed_list[[i]]
 # Perform propensity score stratification or matching 
  ps_strata <- matchit(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
               filtered_imputed_list[[i]],
              method = "nearest",
                    caliper = 0.1, 
               distance = "glm")
  # Add weights from matchit to the original data
imputed_subset$weights <- ps_strata$weights

# Label matched (weights > 0) vs. unmatched
imputed_subset$matched <- ifelse(imputed_subset$weights > 0, 1, 0)

# Group-level summary
match_group_summary <- imputed_subset %>%
  group_by(treatment) %>%  # replace with your group variable name if different
  summarise(
    Original_N = n(),
    Matched_N = sum(matched),
    Excluded_N = sum(matched == 0),
    Prop_Excluded = mean(matched == 0),
    .groups = "drop"
  ) %>%
  mutate(Imputation = i)

matching_summary[[i]] <- match_group_summary
  # summary(ps_strata, subclass = TRUE)
  match.data <- match.data(ps_strata)
  ps_df <- data.frame(
  ePS = ps_strata$distance,
  treatment = imputed_subset$treatment,
  imputation = i
)

# Save to a list or bind all iterations
if (i == 1) {
  all_eps <- ps_df
} else {
  all_eps <- rbind(all_eps, ps_df)
}
  
  table1 <- CreateTableOne(vars = vars, strata = "treatment", data = match.data, test = FALSE, smd = TRUE)
    
  smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
  imputation_results[[i]] <- smd_values
  
  # HR estimates
   cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =  match.data) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
}

all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)


# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()

desired_order <- dataPlotMelt$variable 

# average SMDs across imputations
smd_avg_nn0.1_1<- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))

smd_avg_nn0.1_1$Variable <- factor(smd_avg_nn0.1_1$Variable, levels = desired_order)
smd_avg_nn0.1_1 <- smd_avg_nn0.1_1 %>% arrange(Variable)

all_smds_2 <- all_smds %>% left_join(smd_avg_nn0.1_1, by = "Variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")

all_smds_3$Variable <- fct_recode(all_smds_3$Variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_NNM02_SMD_Prop.png", a, width = 8, height = 6)

library(ggplot2)
all_eps$treatment <- factor(all_eps$treatment, levels = c(0, 1), labels = c("White", "Asian"))

my_plot <- ggplot(all_eps, aes(x = ePS, fill = as.factor(treatment))) +
  geom_density(alpha = 0.4) +
  labs(x = "Estimated Propensity Score", fill = "Race/Ethnicity") +
  theme_minimal()

saveRDS(my_plot, "Asian_density.rds")

my_plot <- ggplot(all_eps, aes(x = as.factor(treatment), y = ePS, fill = as.factor(treatment))) +
  geom_boxplot() +
  labs(x = "Race/Ethnicity", y = "Estimated Propensity Score", fill = "Race/Ethnicity") +
  theme_minimal()

saveRDS(my_plot, "Asian_boxplot.rds")


library(dplyr)

# Mean ePS per imputation and group
avg_eps <- all_eps %>%
  group_by(imputation, treatment) %>%
  summarise(mean_eps = mean(ePS, na.rm = TRUE)) %>%
  group_by(treatment) %>%
  summarise(average_eps = mean(mean_eps))

print(avg_eps)

# HR estimates

# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
nn_0.1_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(nn_0.1_results)


library(dplyr)

match_stats_grouped <- bind_rows(matching_summary)

match_summary_by_group_nnmcaliper <- match_stats_grouped %>%
  group_by(treatment) %>%
  summarise(
    Avg_Original_N = mean(Original_N),
    Avg_Matched_N = mean(Matched_N),
    Avg_Excluded_N = mean(Excluded_N),
    Avg_Prop_Excluded = mean(Prop_Excluded)
  )

print(match_summary_by_group_nnmcaliper)
``` 
```{r}
imputation_results <- list()
coef_list <- list()
se_list <- list()
ps_all <- list()
for (i in 1:22) { 
  # Access the i-th imputed dataset 
 imputed_subset <- filtered_imputed_list[[i]]
 # Perform propensity score stratification or matching 
  ps_strata <- matchit(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
               filtered_imputed_list[[i]],
              method = "nearest",
                    caliper = 0.1, 
               distance = "glm")
  # summary(ps_strata, subclass = TRUE)
  match.data <- match.data(ps_strata)
   ps_df <- data.frame(
    ps = match.data$distance,
    treatment = match.data$treatment,
    imputation = i
)

ps_all[[i]] <- ps_df
  
  table1 <- CreateTableOne(vars = vars, strata = "treatment", data = match.data, test = FALSE, smd = TRUE)
    
  smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
  imputation_results[[i]] <- smd_values
  
  # HR estimates
   cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =  match.data) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
}


ps_combined <- bind_rows(ps_all)
ps_combined$treatment <- factor(ps_combined$treatment, levels = c(0, 1), labels = c("White", "Asian"))


plot_nnm0.1 <-  ggplot(ps_combined, aes(x = ps, fill =as.factor(treatment))) +
geom_density(alpha = 0.6) +
  scale_fill_manual(values = c("White" = "blue", "Asian" = "pink")) +
  labs(x = "PS",
     fill ="Race/Ethnicity") +
theme_minimal() 
print(plot_nnm0.1)
saveRDS(plot_nnm0.1, "Asian_NNM_0.1_PS.rds")
```

# Nerest Neighbor Matching - Without Caliper
```{r}
# #Matching the multiply imputed datasets manually
# match.list <- lapply(1:22, function(i) {
# MatchIt::matchit(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy+sesindex2015_2019:year_diag_group +
#           sesindex2015_2019:issstage + 
#           sesindex2015_2019:practicetype +
#           year_diag_group:age_group +
#           year_diag_group:issstage + 
#           year_diag_group:hemoglobin + 
#           year_diag_group:practicetype +
#           age_group:issstage + 
#           age_group:hemoglobin +
#           age_group:practicetype+
#           issstage:hemoglobin + 
#           issstage:practicetype + 
#           practicetype:hemoglobin,
#                filtered_imputed_list[[i]],
#               method = "nearest",
#                distance = "glm") # method of estimating propensity scores
# })

# create a vector of covariates 
vars <- c( "gender", "age_group", "combined_ecog", "issstage", "asct", "risk_group",
          "sesindex2015_2019","year_diag_group", "hemoglobin","marker_calcium_result",
          "marker_creatinie_result","practicetype",
         "Deletion.13q", "Ploidy")

# # obtain each imputed dataset (interaction terms)
# match.data <- lapply(1:22, function(i) {
#  ps_model <- matchit(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy+sesindex2015_2019:year_diag_group +
#           sesindex2015_2019:issstage + 
#           sesindex2015_2019:practicetype +
#           year_diag_group:age_group +
#           year_diag_group:issstage + 
#           year_diag_group:hemoglobin + 
#           year_diag_group:practicetype +
#           age_group:issstage + 
#           age_group:hemoglobin +
#           age_group:practicetype+
#           issstage:hemoglobin + 
#           issstage:practicetype + 
#           practicetype:hemoglobin,
#                filtered_imputed_list[[i]],
#                method = "nearest",
#                distance = "glm")
#  match.data <- match.data(ps_model)
#  return(match.data)
# })
# 
# # matched SMD
# ## For loop to create and print Table 1 for matched data 
# for (i in 1:22) { 
#   # Generate Table 1 for the i-th imputed dataset 
#   table_matched <- CreateTableOne( vars = vars, strata = "treatment", data =match.data[[i]], test = FALSE, addOverall = TRUE, smd = TRUE) 
#   # Print the iteration number and the Table 1 
#   print(paste("Table 1 for Imputed Dataset:", i)) 
#   print(table_matched, smd = TRUE) # Add SMD if needed 
#   }

```

```{r}
imputation_results <- list()
coef_list <- list()
se_list <- list()
matching_summary <- list()
ps_density_list <- list()

for (i in 1:22) { 
  # Access the i-th imputed dataset 
 imputed_subset <- filtered_imputed_list[[i]]
 # Perform propensity score stratification or matching 
  ps_strata <- matchit(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
               filtered_imputed_list[[i]],
              method = "nearest",
               distance = "glm")
  # Add weights from matchit to the original data
imputed_subset$weights <- ps_strata$weights
# Estimate PS before matching
ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
data = imputed_subset, family = binomial())

imputed_subset$ps <- predict(ps_model, type = "response")
# Store PS + group (e.g., race) for density plotting
ps_density_list[[i]] <- imputed_subset %>%
dplyr::select(ps, treatment) %>% # <-- replace `race_group` with your actual race variable
dplyr::mutate(imputation = i)



# Label matched (weights > 0) vs. unmatched
imputed_subset$matched <- ifelse(imputed_subset$weights > 0, 1, 0)

# Group-level summary
match_group_summary <- imputed_subset %>%
  group_by(treatment) %>%  # replace with your group variable name if different
  summarise(
    Original_N = n(),
    Matched_N = sum(matched),
    Excluded_N = sum(matched == 0),
    Prop_Excluded = mean(matched == 0),
    .groups = "drop"
  ) %>%
  mutate(Imputation = i)

matching_summary[[i]] <- match_group_summary
  # summary(ps_strata, subclass = TRUE)
  match.data <- match.data(ps_strata)
   ps_df <- data.frame(
    ps = match.data$distance,
    treatment = match.data$treatment,
    imputation = i
)

ps_all[[i]] <- ps_df
  
  
  table1 <- CreateTableOne(vars = vars, strata = "treatment", data = match.data, test = FALSE, smd = TRUE)
    
  smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
  imputation_results[[i]] <- smd_values
  
  # HR estimates
   cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =  match.data) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
}

all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)

# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()

# average SMDs across imputations
# tabMatched_nn<- all_smds%>% group_by(Variable) %>%
#   summarise(Mean_SMD = round(mean(X1.vs.2),3)
#             ) %>% arrange(desc(abs(Mean_SMD)))
# print(tabMatched_nn)

smd_avg_nn0.1_1<- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))

smd_avg_nn0.1_1$Variable <- factor(smd_avg_nn0.1_1$Variable, levels = desired_order)
smd_avg_nn0.1_1 <- smd_avg_nn0.1_1 %>% arrange(Variable)

all_smds_2 <- all_smds %>% left_join(smd_avg_nn0.1_1, by = "Variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")

all_smds_3$Variable <- fct_recode(all_smds_3$Variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_NNM_SMD_Prop.png", a, width = 8, height = 6)


#overall
ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))

# HR estimates

# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
nn_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(nn_results)


ps_combined <- bind_rows(ps_all)
ps_combined$treatment <- factor(ps_combined$treatment, levels = c(0, 1), labels = c("White", "Asian"))


plot_nnm <-  ggplot(ps_combined, aes(x = ps, fill =as.factor(treatment))) +
geom_density(alpha = 0.6)  + scale_fill_manual(values = c("White" = "blue", "Asian" = "pink")) +
labs(x = "PS",
     fill ="Race/Ethnicity") +
theme_minimal() + theme(legend.position = "none")
print(plot_nnm)

saveRDS(plot_nnm, "Asian_NNM_PS.rds")

library(dplyr)

match_stats_grouped <- bind_rows(matching_summary)

match_summary_by_group_nnm <- match_stats_grouped %>%
  group_by(treatment) %>%
  summarise(
    Avg_Original_N = mean(Original_N),
    Avg_Matched_N = mean(Matched_N),
    Avg_Excluded_N = mean(Excluded_N),
    Avg_Prop_Excluded = mean(Prop_Excluded)
  )

print(match_summary_by_group_nnm)
ps_density_data <- bind_rows(ps_density_list)
ps_density_data$treatment <- factor(ps_density_data$treatment, levels = c(0, 1), labels = c("White", "Asian"))


p <- ggplot(ps_density_data, aes(x = ps, fill = as.factor(treatment))) +  # Replace with your group var
geom_density(alpha = 0.6)+
  scale_fill_manual(values = c("White" = "blue", "Asian" = "pink")) +labs(x = "PS",
y = "density", fill = "Race/Ethnicity") + theme_minimal() 


saveRDS(p, "Asian_unmatched_ps.rds")
``` 

# Optimal Matching 
```{r}
# create a vector of covariates 
vars <- c( "gender", "age_group", "combined_ecog", "issstage", "asct", "risk_group",
          "sesindex2015_2019","year_diag_group", "hemoglobin","marker_calcium_result",
          "marker_creatinie_result","practicetype",
         "Deletion.13q", "Ploidy")

# # obtain each imputed dataset
# match.data <- lapply(1:22, function(i) {
# ps_model <- matchit(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy+sesindex2015_2019:year_diag_group +
#           sesindex2015_2019:issstage + 
#           sesindex2015_2019:practicetype +
#           year_diag_group:age_group +
#           year_diag_group:issstage + 
#           year_diag_group:hemoglobin + 
#           year_diag_group:practicetype +
#           age_group:issstage + 
#           age_group:hemoglobin +
#           age_group:practicetype+
#           issstage:hemoglobin + 
#           issstage:practicetype + 
#           practicetype:hemoglobin,
#                filtered_imputed_list[[i]],
#                method = "optimal",
#                distance = "glm")
#  match.data <- match.data(ps_model)
#  return(match.data)
# })
# 
# # matched SMD
# ## For loop to create and print Table 1 for matched data 
# for (i in 1:22) { 
#   # Generate Table 1 for the i-th imputed dataset 
#   table_matched <- CreateTableOne( vars = vars, strata = "treatment", data =match.data[[i]], test = FALSE, addOverall = TRUE, smd = TRUE) 
#   # Print the iteration number and the Table 1 
#   print(paste("Table 1 for Imputed Dataset:", i)) 
#   print(table_matched, smd = TRUE) # Add SMD if needed 
#   }

```

```{r}
imputation_results <- list()
matching_summary <- list()
coef_list <- list()
se_list <- list()
for (i in 1:22) { 
  # Access the i-th imputed dataset 
 imputed_subset <- filtered_imputed_list[[i]]
 # Perform propensity score stratification or matching 
  ps_strata <- matchit(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
               filtered_imputed_list[[i]],
              method = "optimal",
               distance = "glm")
  # Add weights from matchit to the original data
imputed_subset$weights <- ps_strata$weights

# Label matched (weights > 0) vs. unmatched
imputed_subset$matched <- ifelse(imputed_subset$weights > 0, 1, 0)

# Group-level summary
match_group_summary <- imputed_subset %>%
  group_by(treatment) %>%  # replace with your group variable name if different
  summarise(
    Original_N = n(),
    Matched_N = sum(matched),
    Excluded_N = sum(matched == 0),
    Prop_Excluded = mean(matched == 0),
    .groups = "drop"
  ) %>%
  mutate(Imputation = i)

matching_summary[[i]] <- match_group_summary
  # summary(ps_strata, subclass = TRUE)
  match.data <- match.data(ps_strata) 
   ps_df <- data.frame(
    ps = match.data$distance,
    treatment = match.data$treatment,
    imputation = i
)

ps_all[[i]] <- ps_df
  
  table1 <- CreateTableOne(vars = vars, strata = "treatment", data = match.data, test = FALSE, smd = TRUE)
    
  smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
  imputation_results[[i]] <- smd_values
  
  # HR estimates
   cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =  match.data) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
}

all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)

# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()

# average SMDs across imputations
tabMatched_optimal_1<- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))
tabMatched_optimal_1$Variable <- factor(tabMatched_optimal_1$Variable, levels = desired_order)
tabMatched_optimal_1 <- tabMatched_optimal_1 %>% arrange(Variable)
print(tabMatched_optimal_1)


all_smds_2 <- all_smds %>% left_join(smd_avg_nn0.1_1, by = "Variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")

all_smds_3$Variable <- fct_recode(all_smds_3$Variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_optimal_SMD_Prop.png", a, width = 8, height = 6)


#overall
ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))

# HR estimates

# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
optimal_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(optimal_results)

#################################################################################################################
# imputation_results <- list()
# for (i in 1:22) { 
#   # Access the i-th imputed dataset 
#  imputed_subset <- filtered_imputed_list[[i]]
#  # Perform propensity score stratification or matching 
#   ps_strata <- matchit(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy
#                + sesindex2015_2019:year_diag_group+
#           sesindex2015_2019:issstage + 
#           sesindex2015_2019:practicetype +
#           sesindex2015_2019:combined_ecog +
#             year_diag_group:issstage +
#             year_diag_group:practicetype +
#             year_diag_group:combined_ecog +
#             issstage:practicetype +
#             issstage:combined_ecog + 
#             practicetype:combined_ecog,
#                imputed_subset,
#                method = "optimal",
#                distance = "glm")
#   # summary(ps_strata, subclass = TRUE)
#   match.data <- match.data(ps_strata)
# 
#     table1 <- CreateTableOne(vars = vars, strata = "treatment", data =  match.data, test = FALSE, smd = TRUE)
#     
#     smd_values <- data.frame(
#       Variable = vars,
#       SMD = ExtractSmd(table1),
#       Imputation = i,
#       stringsAsFactors = FALSE)
#     
#   imputation_results[[i]] <- smd_values
# }
# all_smds <- do.call(rbind, imputation_results)
# 
# # generating proprotions 
# library(dplyr)
# proportions_by_variable <- all_smds %>%
#   group_by(Variable) %>%
#   summarise(
#     proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
#   )
# 
# print(proportions_by_variable)
# 
# # visualization
# ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
#                                                y =proportion_exceeding_0.1)) +
#   geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
#                                                          x = "Covariate",
#                                                          y = "Proportions Exceeding 0.1") +
#   theme_minimal() + coord_flip()
# 
# # average SMDs across imputations
# tabMatched_optimal <- all_smds%>% group_by(Variable) %>%
#   summarise(Mean_SMD = round(mean(X1.vs.2),3)
#             ) %>% arrange(desc(abs(Mean_SMD)))
# print(tabMatched_optimal)
# 
# all_smds_2 <- all_smds %>% left_join(tabMatched_optimal, by = "Variable")
# all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")
# 
# #overall
# ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
#   geom_point(alpha=0.5, size =2) +
#   geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
#   geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))


ps_combined <- bind_rows(ps_all)
ps_combined$treatment <- factor(ps_combined$treatment, levels = c(0, 1), labels = c("White", "Asian"))


plot_optimal<-  ggplot(ps_combined, aes(x = ps, fill =as.factor(treatment))) +
geom_density(alpha = 0.6)  + scale_fill_manual(values = c("White" = "blue", "Asian" = "pink")) +
labs(x = "PS",
     fill ="Race/Ethnicity") +
theme_minimal() + theme(legend.position = "none")
print(plot_optimal)
saveRDS(plot_optimal, "Asian_optimal_PS.rds")
library(dplyr)

match_stats_grouped <- bind_rows(matching_summary)

match_summary_by_group_optimal <- match_stats_grouped %>%
  group_by(treatment) %>%
  summarise(
    Avg_Original_N = mean(Original_N),
    Avg_Matched_N = mean(Matched_N),
    Avg_Excluded_N = mean(Excluded_N),
    Avg_Prop_Excluded = mean(Prop_Excluded)
  )

print(match_summary_by_group_optimal)
``` 


#IPTW (without trimming) - for SMD calculations as it is the same as stabilized iptw
```{r}
library(survey)
imputation_results <- list()
for (i in 1:22) { 
  # Access the i-th imputed dataset 
 dat <- filtered_imputed_list[[i]]
dat$gender <- as.factor(dat$gender)
dat$treatment<- as.factor(dat$treatment)
ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
              dat, family = binomial())

dat$pscore <- predict(ps_model, type="response")
dat$ps_weights <- ifelse(dat$treatment == 1, 1/dat$pscore, 1/(1-dat$pscore))
design <- svydesign(ids = ~1, data =dat, weights = ~ps_weights)

    table1 <- svyCreateTableOne( vars =  vars, strata = "treatment", data = design, test =FALSE, addOverall = TRUE)
    
    smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
    print(summary(dat$ps_weights))
    plot(dat$ps_weights)
  imputation_results[[i]] <- smd_values
}
all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)

# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()

# average SMDs across imputations
smd_avg_iptw_1 <- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))
smd_avg_iptw_1$Variable <- factor(smd_avg_iptw_1$Variable, levels = desired_order)
smd_avg_iptw_1 <- smd_avg_iptw_1 %>% arrange(Variable)
print(smd_avg_iptw_1)

all_smds_2 <- all_smds %>% left_join(smd_avg_iptw_1, by = "Variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")

all_smds_3$Variable <- fct_recode(all_smds_3$Variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_SIPTW_SMD_Prop.png", a, width = 8, height = 6)


#overall
ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))
################################################################################################################
# imputation_results <- list()
# for (i in 1:22) { 
#   # Access the i-th imputed dataset 
#  dat <- filtered_imputed_list[[i]]
# dat$gender <- as.factor(dat$gender)
# dat$treatment<- as.factor(dat$treatment)
# ps_model <- glm(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy + 
#                  sesindex2015_2019:gender,
#               dat, family = binomial())
# # wt<-WeightIt::weightit(treatment ~ age_group+
# #                gender + issstage + asct + combined_ecog+
# #                  risk_group + sesindex2015_2019 +
# #                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
# #                + practicetype + Deletion.13q + Ploidy,
# #               dat,
# #               stabilize = TRUE,
# #               method = "ps")
# # dat$ps_weights <- wt$weights
# dat$pscore <- predict(ps_model, type="response")
# dat$ps_weights <- ifelse(dat$treatment == 1, 1/dat$pscore, 1/(1-dat$pscore))
# design <- svydesign(ids = ~1, data =dat, weights = ~ps_weights)
# 
#     table1 <- svyCreateTableOne( vars =  vars, strata = "treatment", data = design, test =FALSE, addOverall = TRUE)
#     
#     smd_values <- data.frame(
#       Variable = vars,
#       SMD = ExtractSmd(table1),
#       Imputation = i,
#       stringsAsFactors = FALSE)
#     
#     print(summary(dat$ps_weights))
#     plot(dat$ps_weights)
#   imputation_results[[i]] <- smd_values
# }
# all_smds <- do.call(rbind, imputation_results)
# 
# # generating proprotions 
# library(dplyr)
# proportions_by_variable <- all_smds %>%
#   group_by(Variable) %>%
#   summarise(
#     proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
#   )
# 
# print(proportions_by_variable)
# 
# # visualization
# ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
#                                                y =proportion_exceeding_0.1)) +
#   geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
#                                                          x = "Covariate",
#                                                          y = "Proportions Exceeding 0.1") +
#   theme_minimal() + coord_flip()
# 
# # average SMDs across imputations
# smd_avg_iptw <- all_smds%>% group_by(Variable) %>%
#   summarise(Mean_SMD = round(mean(X1.vs.2),3)
#             ) %>% arrange(desc(abs(Mean_SMD)))
# print(smd_avg_iptw)
# 
# all_smds_2 <- all_smds %>% left_join(smd_avg_iptw, by = "Variable")
# all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")
# 
# #overall
# ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
#   geom_point(alpha=0.5, size =2) +
#   geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
#   geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))
```

#IPTW (with trimming)

```{r}
imputation_results <- list()
coef_list <- list()
se_list <- list()
trim_summary <- list()
for (i in 1:22) { 
  dat <- filtered_imputed_list[[i]]
  dat$gender <- as.factor(dat$gender)
  dat$treatment<- as.factor(dat$treatment)
  ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
              dat, family = binomial())
  dat$pscore <- predict(ps_model, type="response")
  dat$ps_weights_2 <- ifelse(dat$treatment == 1, 1/dat$pscore, 1/(1-dat$pscore))

  # apply trimming to weights
  lower_threshold <- quantile(dat$ps_weights_2 ,0.05)
  upper_threshold <- quantile(dat$ps_weights_2 ,0.95)
  dat$trimmed_out <- !(dat$ps_weights_2 >= lower_threshold & dat$ps_weights_2 <= upper_threshold)
  aa_data_trimmed <- dat %>% filter(ps_weights_2  >=lower_threshold & 
                                        ps_weights_2  <= upper_threshold)
  # Group-level summary of trimming
trim_group_summary <- dat %>%
  group_by(treatment) %>%  # replace 'treatment' if your variable is named differently
  summarise(
    Original_N = n(),
    Trimmed_N = sum(trimmed_out),
    Retained_N = n() - sum(trimmed_out),
    Prop_Trimmed = mean(trimmed_out),
    .groups = "drop"
  ) %>%
  mutate(Imputation = i)

trim_summary[[i]] <- trim_group_summary

  design <- svydesign(ids = ~1, data =aa_data_trimmed , weights = ~ps_weights_2 )

  table1 <- svyCreateTableOne( vars =  vars, strata = "treatment", data = design, test =FALSE, addOverall = TRUE)
    
  smd_values <- data.frame(
      Variable = vars,
      SMD = ExtractSmd(table1),
      Imputation = i,
      stringsAsFactors = FALSE)
    
  imputation_results[[i]] <- smd_values
  
  # HR estimates
  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =aa_data_trimmed, weights = ps_weights_2) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
  
    all_weights_df[[i]] <- data.frame(
    id = 1:nrow(aa_data_trimmed),
    weight = aa_data_trimmed$ps_weights_2,
    treatment = aa_data_trimmed$treatment)
    
  
}

library(dplyr)

trim_stats_grouped <- bind_rows(trim_summary)

trim_summary_by_group <- trim_stats_grouped %>%
  group_by(treatment) %>%
  summarise(
    Avg_Original_N = mean(Original_N),
    Avg_Retained_N = mean(Retained_N),
    Avg_Trimmed_N = mean(Trimmed_N),
    Avg_Prop_Trimmed = mean(Prop_Trimmed)
  )

print(trim_summary_by_group)

all_smds <- do.call(rbind, imputation_results)

# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(Variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
  )

print(proportions_by_variable)

# visualization
ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
                                               y =proportion_exceeding_0.1)) +
  geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
                                                         x = "Covariate",
                                                         y = "Proportions Exceeding 0.1") +
  theme_minimal() + coord_flip()

# average SMDs across imputations
iptw_trimmed_1 <- all_smds%>% group_by(Variable) %>%
  summarise(Mean_SMD = round(mean(X1.vs.2),3)
            ) %>% arrange(desc(abs(Mean_SMD)))
iptw_trimmed_1$Variable <- factor(iptw_trimmed_1$Variable, levels = desired_order)
iptw_trimmed_1  <- iptw_trimmed_1  %>% arrange(Variable)
print(iptw_trimmed_1)

all_smds_2 <- all_smds %>% left_join(iptw_trimmed_1, by = "Variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")

all_smds_3$Variable <- fct_recode(all_smds_3$Variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_IPTWTRIM_SMD_Prop.png", a, width = 8, height = 6)


#overall
ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))

# HR estimates

# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
iptw_trimming_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(iptw_trimming_results)

# Bind all imputations together
combined_df <- bind_rows(all_weights_df, .id = "imp")

# Compute average weight per individual across imputations
avg_weights_df <- combined_df %>%
  group_by(id, treatment) %>%
  summarise(avg_weight = mean(weight, na.rm = TRUE), .groups = "drop")
avg_weights_df$treatment <- factor(avg_weights_df $treatment, levels = c(0, 1), labels = c("White", "Asian"))

ggplot(avg_weights_df, aes(x = as.factor(treatment), y = avg_weight)) +
  geom_boxplot(fill = "pink") +
  labs(
       x = "Race/Ethnicity Group",
       y = "Average Stablized Weight") +theme_minimal()

plot_trimming_iptw <- ggplot(avg_weights_df, aes(x = avg_weight)) +
  geom_density(alpha = 0.6,fill ="blue") +
  labs(
       x = "IPTW Weight", y = "Frequency") + theme_minimal() + theme(legend.position = "none")

saveRDS(plot_trimming_iptw, "Asian_trimming.rds")

```

#IPTW (stablized)
```{r}
imputation_results <- list()
coef_list <- list()
se_list <- list()
all_weights_df <- list()

for (i in 1:22) { 
  # Access the i-th imputed dataset 
 imputed_subset <- filtered_imputed_list[[i]]
 # Perform propensity score stratification or matching 
 ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
              imputed_subset, family = binomial())
 imputed_subset$pscore <- predict(ps_model, type="response")
 p_exposure <- mean(imputed_subset$treatment==1)
 p_non_exposure <- 1-p_exposure

 imputed_subset$ps_weights_3 <- ifelse(imputed_subset$treatment == 1, 
                           p_exposure/imputed_subset$pscore, 
                           p_non_exposure/(1-imputed_subset$pscore))
  # HR estimates
  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data =  imputed_subset, weights = ps_weights_3) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
  
  
  all_weights_df[[i]] <- data.frame(
    id = 1:nrow(imputed_subset),
    weight = imputed_subset$ps_weights_3,
    treatment = imputed_subset$treatment
  )
  
}



# Bind all imputations together
combined_df <- bind_rows(all_weights_df, .id = "imp")

# Compute average weight per individual across imputations
avg_weights_df <- combined_df %>%
  group_by(id, treatment) %>%
  summarise(avg_weight = mean(weight, na.rm = TRUE), .groups = "drop")
avg_weights_df$treatment <- factor(avg_weights_df $treatment, levels = c(0, 1), labels = c("White", "Asian"))

ggplot(avg_weights_df, aes(x = as.factor(treatment), y = avg_weight)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Average Stabilized IPTW Weights by Race/Ethnicity",
       x = "Treatment Group",
       y = "Average Weight") +theme_minimal()

plot_sta_iptw <- ggplot(avg_weights_df, aes(x = avg_weight, fill =as.factor(treatment))) +
  geom_density(alpha = 0.6) + scale_fill_manual(values = c("White" = "blue", "Asian" = "pink")) + 
  labs(title = "Histogram of Average Stabilized IPTW Weights by Race/Ethnicity",
       x = "IPTW Stabilized Weight", y = "Frequency", fill ="Race/Ethnicity") + theme_minimal()+ theme(legend.position = "none") 

saveRDS(plot_sta_iptw, "Asian_sta_iptw.rds")


# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
stablized_iptw_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(stablized_iptw_results)

```

# PS as an Additional Covariate
```{r}
# Assuming you have multiple imputed datasets in a list called `imputed_list` 
num_imputations <- length(filtered_imputed_list) 
# Create placeholders to store coefficients and standard errors 
coef_list <- list() 
se_list <- list() 
# Fit Cox model on each imputed dataset 
for (i in 1:num_imputations) { 
  ps_model <- glm(treatment ~ age_group +
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
             filtered_imputed_list[[i]], family = binomial())
  filtered_imputed_list[[i]]$pscore <- predict(ps_model, type="response")

  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment) + age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy +pscore, data = filtered_imputed_list[[i]]) 
  coef_list[[i]] <- coef(cox_model) # Store coefficients 
  se_list[[i]] <- sqrt(diag(vcov(cox_model))) # Store standard errors 
  } 
# Convert lists to matrices 
coef_matrix <- do.call(rbind, coef_list) 
se_matrix <- do.call(rbind, se_list) 
# Calculate pooled estimates (Rubin's Rules) 
mean_coefs <- colMeans(coef_matrix) # Average of the coefficients 
# Calculate within-imputation variance 
within_var <- colMeans(se_matrix^2) 
# Calculate between-imputation variance 
between_var <- apply(coef_matrix, 2, var) 
# Compute total variance using Rubin's formula 
total_var <- within_var + (1 + (1/num_imputations)) * between_var 
# Compute pooled standard errors 
pooled_se <- sqrt(total_var) 
# Compute 95% confidence intervals 
lower_CI <- mean_coefs - 1.96 * pooled_se 
upper_CI <- mean_coefs + 1.96 * pooled_se 
# Convert coefficients to hazard ratios 
hr_estimates <- exp(mean_coefs) 
hr_lower_CI <- exp(lower_CI) 
hr_upper_CI <- exp(upper_CI) 
# Print results 
doubly_robust_results <- data.frame( Variable = names(mean_coefs), HR = hr_estimates, Lower_CI = hr_lower_CI, Upper_CI = hr_upper_CI, se =pooled_se , loghr =mean_coefs  )
print(doubly_robust_results)
```

# Overlap Weighting
```{r}
# #install.packages("PSweight")
# library(PSweight)
# imputation_results <- list()
# for (i in 1:22) { 
# dat <- filtered_imputed_list[[i]]
# dat$gender <- as.factor(dat$gender)
# dat$treatment<- as.factor(dat$treatment)
# ps_model <- glm(treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy,
#               dat, family = binomial())
# # wt<-WeightIt::weightit(treatment ~ age_group+
# #                gender + issstage + asct + combined_ecog+
# #                  risk_group + sesindex2015_2019 +
# #                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
# #                + practicetype + Deletion.13q + Ploidy,
# #               dat,
# #               stabilize = TRUE,
# #               method = "ps")
# # dat$ps_weights <- wt$weights
# dat$pscore <- predict(ps_model, type="response")
# dat$overlap_weights <- ifelse(dat$treatment == 1, 1-dat$pscore, dat$pscore)
# 
# 
# design <- svydesign(ids = ~1, data =dat, weights = ~overlap_weights )
# 
#     table1 <- svyCreateTableOne( vars =  vars, strata = "treatment", data = design, test =FALSE, addOverall = TRUE)
#     
#     smd_values <- data.frame(
#       Variable = vars,
#       SMD = ExtractSmd(table1),
#       Imputation = i,
#       stringsAsFactors = FALSE)
#     
#     print(summary(dat$overlap_weights))
#     plot(dat$overlap_weights)
#   imputation_results[[i]] <- smd_values
# }
# all_smds <- do.call(rbind, imputation_results)
# 
# # generating proprotions 
# library(dplyr)
# proportions_by_variable <- all_smds %>%
#   group_by(Variable) %>%
#   summarise(
#     proportion_exceeding_0.1 = sum(X1.vs.2 >0.1)/n()
#   )
# print(proportions_by_variable)
# 
# # visualization
# ggplot(proportions_by_variable , aes(x=reorder(Variable,proportion_exceeding_0.1),
#                                                y =proportion_exceeding_0.1)) +
#   geom_bar(stat = "identity", fill = "steelblue") + labs(title = "Proportions of SMDs Exceeding 0.1 by Covariate",
#                                                          x = "Covariate",
#                                                          y = "Proportions Exceeding 0.1") +
#   theme_minimal() + coord_flip()
# 
# # average SMDs across imputations
# smd_avg_iptw_1 <- all_smds%>% group_by(Variable) %>%
#   summarise(Mean_SMD = round(mean(X1.vs.2),3)
#             ) %>% arrange(desc(abs(Mean_SMD)))
# print(smd_avg_iptw_1)
# 
# all_smds_2 <- all_smds %>% left_join(smd_avg_iptw_1, by = "Variable")
# all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "Variable")
# 
# ggplot(dat, aes(x = pscore, fill = factor(treatment)))+
#   geom_density(alpha = 0.5) 
# 
# #overall
# ggplot(all_smds_3, aes(x = X1.vs.2, y = Variable)) + 
#   geom_point(alpha=0.5, size =2) +
#   geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
#   geom_point( aes(x = Mean_SMD, y = Variable), color ="blue", size = 3) +  labs(title = "SMDs Across Imputations",x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(all_smds_3$X1.vs.2+0.05), y = reorder(Variable,proportion_exceeding_0.1),  label =  round( proportion_exceeding_0.1,2)))
# 
# cox_model1 <- svycoxph(Surv(OS, event)  ~ treatment, design = design)
# print(cox_model1)
# 
# dat$Y <- Surv(dat$OS, dat$event)
# 
# a<- PSweight(
#   ps.formula = treatment ~ age_group+
#                gender + issstage + asct + combined_ecog+
#                  risk_group + sesindex2015_2019 +
#                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
#                + practicetype + Deletion.13q + Ploidy,
#   yname = 'Y',
#   data=dat,
#   weight = 'overlap',
#   out.method = "coxph",
# )
# a
# 
# boxplot(overlap_weights ~ treatment, data = dat)
# ggplot(dat, aes(x = overlap_weights, fill = factor(treatment))) +
#   geom_density(alpha = 0.5)
# 
# # library(cobalt)
# # bal_tab <- bal.tab(treatment ~ age_group+
# #                gender + issstage + asct + combined_ecog+
# #                  risk_group + sesindex2015_2019 +
# #                  year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
# #                + practicetype + Deletion.13q + Ploidy,
# #                data = dat,
# #                weights = dat$overlap_weights)
# # love.plot(bal_tab)
# 
# print(table1, smd = TRUE, digits = 10)
# table2 <- CreateTableOne( vars =  vars, strata = "treatment", data = dat, test =FALSE, addOverall = TRUE)
# print(table2, smd = TRUE)
# ```
# ##### Comparisons of Covariate Balance
# ```{r}
# ## Construct a data frame containing variable name and SMD from all methods
# dataPlot <- data.frame(variable   = tabUnmatched$Variable,
#                        Unmatched  = as.numeric(tabUnmatched$Mean_SMD),
#                       
#                        Matched_caliper_0.1  = as.numeric(smd_avg_nn0.1_1$Mean_SMD),
#                        Matched_nerest_neighbor = as.numeric(tabMatched_caliper_1$Mean_SMD),
#                        IPTW = as.numeric(smd_avg_iptw_1$Mean_SMD),
#                        IPTW_with_trimming = as.numeric(iptw_trimmed_1$Mean_SMD)
#                        # weighted_iptw =as.numeric(ExtractSmd(tabWeighted))
#                        )
# # Matched_optimal  = as.numeric(tabMatched_optimal_1$Mean_SMD),
# ## Create long-format data for ggplot2
# dataPlotMelt <- melt(data          = dataPlot,
#                      id.vars       = c("variable"),
#                      variable.name = "Method",
#                      value.name    = "SMD")
# 
# ## Order variable names by magnitude of SMD
# varNames <- as.character(dataPlot$variable)[order(dataPlot$Unmatched,decreasing = FALSE)]
# 
# ## Order factor levels in the same order
# dataPlotMelt$variable <- factor(dataPlotMelt$variable,
#                                 levels = varNames)
# 
# ## Plot using ggplot2
# ggplot(data = dataPlotMelt,
#        mapping = aes(x = variable, y = SMD, group = Method, color = Method)) +
#     geom_line() +
#     geom_point() +
#     geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
#     coord_flip() +
#     theme_bw() +
#     theme(legend.key = element_blank()) + scale_y_continuous(limits = c(0,0.5))
# ###############################################################################################################
# ## Construct a data frame containing variable name and SMD from all methods
# dataPlot <- data.frame(variable   = tabUnmatched$Variable,
#                        Unmatched  = as.numeric(tabUnmatched$Mean_SMD),
#                       
#                        Matched_caliper_0.1  = as.numeric(smd_avg_nn0.1$Mean_SMD),
#                        Matched_nerest_neighbor = as.numeric(tabMatched_caliper$Mean_SMD),
#                        IPTW = as.numeric(smd_avg_iptw$Mean_SMD),
#                        IPTW_with_trimming = as.numeric(iptw_trimmed$Mean_SMD)
#                        # weighted_iptw =as.numeric(ExtractSmd(tabWeighted))
#                        )
# # Matched_optimal  = as.numeric(tabMatched_optimal$Mean_SMD),
# 
# ## Create long-format data for ggplot2
# dataPlotMelt <- melt(data          = dataPlot,
#                      id.vars       = c("variable"),
#                      variable.name = "Method",
#                      value.name    = "SMD")
# 
# ## Order variable names by magnitude of SMD
# varNames <- as.character(dataPlot$variable)[order(dataPlot$Unmatched,decreasing = FALSE)]
# 
# ## Order factor levels in the same order
# dataPlotMelt$variable <- factor(dataPlotMelt$variable,
#                                 levels = varNames)
# 
# ## Plot using ggplot2
# ggplot(data = dataPlotMelt,
#        mapping = aes(x = variable, y = SMD, group = Method, color = Method)) +
#     geom_line() +
#     geom_point() +
#     geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
#     coord_flip() +
#     theme_bw() +
#     theme(legend.key = element_blank()) + scale_y_continuous(limits = c(0,0.5))
```


# PS stratification

```{r}
num_imputations <- length(filtered_imputed_list) 
ate_results <- list() 
att_results <- list()
ps_by_stratum_group_all <- list()

for (i in 1:num_imputations) { 
  imputed_subset <- filtered_imputed_list[[i]] 
  
  # Step 1: Estimate propensity scores 
  ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
              imputed_subset, family = binomial())
  
  imputed_subset$ps <- predict(ps_model, type = "response") 
  
  # Step 2: Create 5 strata based on quintiles 
  imputed_subset$strata <- cut(imputed_subset$ps, breaks = quantile(imputed_subset$ps, probs = seq(0, 1, 0.2), na.rm = TRUE), include.lowest = TRUE, labels = FALSE) 
# Summarize by stratum and group
ps_by_stratum_group_all[[i]] <- imputed_subset %>%
dplyr::group_by(strata, treatment) %>% # Replace `group` with your variable, e.g., race
dplyr::summarize(mean_ps = mean(ps, na.rm = TRUE), .groups = 'drop') %>%
dplyr::mutate(imputation = i)

  # Step 3: Calculate strata weights 
  strata_counts <- imputed_subset %>% 
    group_by(strata) %>% 
    summarise(n_treated  = sum(treatment == 1), 
              n_control = sum(treatment == 0), 
              n_total = n()) 
  
  # Step 4: Fit Cox models within each stratum and extract HR and SE 
  stratum_results <- list() 
  
  for (s in unique(imputed_subset$strata)) { 
    stratum_data <- imputed_subset %>% filter(strata == s) 
    
    # Cox model for ATE (all subjects) 
    cox_model <-coxph(Surv(OS, event) ~ as.factor(treatment), data = stratum_data) 
    
    # Extract log HR and SE
    hr <- exp(coef(cox_model)[1])
    se <- sqrt(diag(vcov(cox_model))[1])
    
    stratum_results[[s]] <- data.frame(
      Stratum = s, 
      HR = hr,
      logHR = log(hr),
      SE = se,
      Treated = strata_counts$n_treated[s],
      Total = strata_counts$n_total[s]
      
    )

  }
    stratum_df <- do.call(rbind, stratum_results)
    # Step 5: Calculate pooled ATE and ATT using stratum sample sizes 
    
    # ATE: Weighted by total stratum size (treated + control) 
    ate_log_hr <- sum(stratum_df$Total * stratum_df$logHR) / sum(stratum_df$Total) 
    ate_var <- sum((stratum_df$Total^2 * stratum_df$SE^2)) / (sum(stratum_df$Total)^2) 
    ate_se <- sqrt(ate_var) 
    
    # ATT: Weighted by number of treated units 
    att_log_hr <- sum(stratum_df$Treated * stratum_df$logHR) / sum(stratum_df$Treated) 
    att_var <- sum((stratum_df$Treated^2 * stratum_df$SE^2)) / (sum(stratum_df$Treated)^2) 
    att_se <- sqrt(att_var) 
    
    # Store pooled results for each imputation 
    ate_results[[i]] <- data.frame(LogHR = ate_log_hr, SE = ate_se) 
    att_results[[i]] <- data.frame(LogHR = att_log_hr, SE = att_se) 
    
    imputed_subset$treatment <- factor( imputed_subset$treatment, levels = c(0, 1), labels = c("White", "Asian"))

     my_plot <- ggplot(imputed_subset, aes(x = as.factor(strata), y = ps, fill = as.factor(treatment))) +
        geom_boxplot() + labs(x = "Stratum", y = "Propensity Score") + theme_minimal()+
        labs(x = "Stratum", fill = "Race/Ethnicity", y = "Estimated Propensity Score")
     saveRDS(my_plot, "Asian_strata.rds")

        } 




# Step 6: Pool results across multiple imputations using Rubin's Rules 
pool_rubin <- function(results_list) { 
  m <- length(results_list) 
  logHR_vals <- sapply(results_list, function(x) x$LogHR) 
  se_vals <- sapply(results_list, function(x) x$SE) 
  
  # Calculate pooled log HR
  pooled_logHR <- mean(logHR_vals) 
  
  # Calculate within-imputation variance 
  within_var <- mean(se_vals^2) 
  
  # Calculate between-imputation variance 
  between_var <- var(logHR_vals) 
  
  # Total variance 
  total_var <- within_var + (1 + (1 / m)) * between_var 
  
  # Final pooled SE 
  pooled_se <- sqrt(total_var) 
  
  # Convert log HR back to HR 
  pooled_hr <- exp(pooled_logHR) 
  lower_ci <- exp(pooled_logHR - 1.96 * pooled_se) 
  upper_ci <- exp(pooled_logHR + 1.96 * pooled_se) 
  
  return(data.frame( Pooled_HR = pooled_hr, SE = pooled_se, Lower_CI = lower_ci, Upper_CI = upper_ci, se =pooled_se , loghr =pooled_logHR )) } 

# Step 7: Get final pooled results for ATE and ATT 
final_ate_results <- pool_rubin(ate_results) 
final_att_results <- pool_rubin(att_results) 

# Print results 
cat("Pooled ATE Results:\n") 
print(final_ate_results) 

cat("\nPooled ATT Results:\n") 
print(final_att_results)

final_ate_results <- final_ate_results %>% rename(HR = Pooled_HR) %>% mutate(Variable = "as.factor(treatment)1")
final_att_results <- final_att_results %>% rename(HR = Pooled_HR) %>% mutate(Variable = "as.factor(treatment)1")

library(ggplot2)

ps_summary_group <- dplyr::bind_rows(ps_by_stratum_group_all)

plot <- ggplot(ps_summary_group, aes(x = factor(strata), y = mean_ps, fill = as.factor(treatment))) +
geom_boxplot(position = position_dodge(width = 0.8)) +
labs(
x = "Stratum",
y = "PS",
fill = "Race/Ethnicity")  +scale_fill_manual(values = c("0" = "blue", "1" = "pink"),
labels = c("White", "Asian")) + theme_minimal()

saveRDS(plot, "Asian_PS stratification.rds")

```

```{r}
# check PH assumptions
aa_data <- filtered_imputed_list[[22]]
aa_data <- aa_data %>% mutate(
  stratum = ntile(pscore,5)
)


for (s in unique(aa_data$stratum)) {
  stratum_data <- subset(aa_data, stratum ==s)
  cox_model <- coxph(Surv(OS, event) ~ as.factor(treatment), data = stratum_data)
  print(cox.zph(cox_model))
}

# covariate balance in each stratum

stratum_smds <- data.frame(stratum = numeric(),
                           covariate = character(),
                           smd = numeric(),
                           n = numeric())
for (s in unique(aa_data$stratum)) {
  stratum_data <- subset(aa_data, stratum ==s)
  table1 <- CreateTableOne(vars = vars, strata = "treatment", data = stratum_data)
  smd<- ExtractSmd(table1)
  stratum_smds <- rbind(stratum_smds,
                        data.frame(stratum = s,
                                   variable = vars,
                                   smd = smd,
                                   n = nrow(stratum_data)
                                   ))
  }

print(stratum_smds)

weighted_smds <- stratum_smds %>% group_by(variable) %>%
  summarise(weighted_smd = sum(`X1.vs.2`*n)/sum(n))

weighted_smds

weighted_smds$variable <- factor(weighted_smds$variable)
weighted_smds<- weighted_smds  %>% arrange(variable)


##########

```

```{r}
imputation_results <- list()

for (i in 1:22) {
cat("Processing imputation", i, "\n")

aa_data <- filtered_imputed_list[[i]]
# Estimate PS before matching
ps_model <- glm(treatment ~ age_group+
               gender + issstage + asct + combined_ecog+
                 risk_group + sesindex2015_2019 +
                 year_diag_group + hemoglobin + marker_calcium_result + marker_creatinie_result
               + practicetype + Deletion.13q + Ploidy,
data = aa_data, family = binomial())

aa_data$ps <- predict(ps_model, type = "response")
aa_data <- aa_data %>%
mutate(stratum = ntile(ps, 5))
stratum_smds <- data.frame()

for (s in unique(aa_data$stratum)) {
stratum_data <- subset(aa_data, stratum == s)

smd_data <- ExtractSmd(CreateTableOne(vars = vars, strata = "treatment", data = stratum_data, test = FALSE))

# Build stratum-level dataframe
stratum_smds <- rbind(stratum_smds,
data.frame(
stratum = s,
variable = vars,
smd = as.numeric(smd_data),
n = nrow(stratum_data)
))
}

# Calculate weighted SMD per variable (NOT collapsed across variables!)
weighted_smds <- stratum_smds %>%
group_by(variable) %>%
summarise(weighted_smd = sum(smd * n) / sum(n), .groups = 'drop')

imputation_results[[i]] <- weighted_smds
}

# Combine all imputations
all_smds <- bind_rows(imputation_results, .id = "imputation_id")

# Pool across imputations (Rubins averaging)
pooled_smds <- all_smds %>%
group_by(variable) %>%
summarise(
mean_smd = mean(weighted_smd, na.rm = TRUE),
sd_smd = sd(weighted_smd, na.rm = TRUE),
n_imputations = n()
)

print(pooled_smds)

weighted_smds <- pooled_smds
weighted_smds  <- weighted_smds %>% rename(weighted_smd = mean_smd)
desired_order <- dataPlotMelt$variable 

weighted_smds$variable <- factor(weighted_smds$variable, levels = tabUnmatched$Variable)
weighted_smds<- weighted_smds  %>% arrange(variable)


# generating proprotions 
library(dplyr)
proportions_by_variable <- all_smds %>%
  group_by(variable) %>%
  summarise(
    proportion_exceeding_0.1 = sum(weighted_smd >0.1)/n()
  )

print(proportions_by_variable)


all_smds_2 <- all_smds %>% left_join(weighted_smds, by = "variable")
all_smds_3 <- all_smds_2 %>% left_join(proportions_by_variable, by = "variable")

            
all_smds_3$variable <- fct_recode(all_smds_3$variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

#overall
a <- ggplot(all_smds_3, aes(x = weighted_smd.x, y = variable)) + 
  geom_point(alpha=0.5, size =2) +
  geom_vline(xintercept = c(0.1), linetype = "dashed", color="red") +
  geom_point( aes(x = weighted_smd.y, y = variable), color ="blue", size = 3) +  labs(x="SMD",y = "Coavariate") + theme_minimal() + geom_text(aes(x = max(weighted_smd.x+0.05), y = reorder(variable, proportion_exceeding_0.1),  label =  round(proportion_exceeding_0.1,2)))

# saveRDS(all_smds_3, "all_smds_3_asian_NNM02.rds")
# saveRDS(a, "Asian_NNM02_SMD_Prop.rds")
ggsave("Asian_STRATA_SMD_Prop.png", a, width = 8, height = 6)

```



# Tables
```{r}
library(dplyr) 
library(purrr) 

# Function to extract and combine treatment effect results 
aggregate_treatment_results <- function(results_list, method_names, treatment_var = "as.factor(treatment)1") 
{ 
  # Add method names to each dataframe 
  named_results <- map2(results_list, method_names, ~mutate(.x, Method = .y)) 
  # Combine all results into one dataframe 
  combined_results <- bind_rows(named_results) 
  # Filter for the treatment variable only 
  treatment_results <- combined_results %>% 
    filter(Variable == treatment_var) %>% select(Method, Variable, HR, Lower_CI, Upper_CI, se, loghr) %>% rename( Hazard_Ratio = HR)
  return(treatment_results) }

```

```{r}
# List of results from different PS methods 
results_list <- list(crude_results, 
                     cov_adj_results, 
                     nn_results,
                     nn_0.1_results,
                     optimal_results,
              #       iptw_trimming_results,
                     stablized_iptw_results,
                   #  final_att_results,
                     final_ate_results,
                     doubly_robust_results
                     ) 
method_names <- c("Crude", 
                  "Covariate Adjustment", 
                  "NNM",
                  "NNM (caliper width = 0.2)",
                  "Optimal Matching",
            #      "IPTW (5% trimming)",
                  "Stablized IPTW",
                 #"PS Stratification (5 strata) - ATT",
                   "PS Stratification (5 strata)",
                  "PS as Covariate"
                  )

c <- aggregate_treatment_results(results_list, method_names)
c <- c %>% select(Method,Hazard_Ratio, Lower_CI, Upper_CI,se, loghr);
rownames(c) <- NULL

b <- c
b$Hazard_Ratio <- round(b$Hazard_Ratio,3)
b$Lower_CI <- round(b$Lower_CI,3)
b$Upper_CI <- round(b$Upper_CI,3)
b$se <- round(b$se,3)
b$loghr <- round(b$loghr,3)
b <- b %>%
  rename(
    HR = Hazard_Ratio,
    `95% CI Lower` = Lower_CI,
    `95% CI Upper` = Upper_CI
  )

print(b)
knitr::kable(b, caption = "Hazard Ratios and 95% Confidence Intervals by Method")
saveRDS(b, "Asian_table.rds")
```

# Forest Plot
```{r}
# Define the desired order manually
desired_order <- c("PS as Covariate",
                   "PS Stratification (5 strata)",
                 #  "PS Stratification (5 strata) - ATT",
                   "Stablized IPTW",
                   "IPTW (5% trimming)",
                    "Optimal Matching",
                   "NNM (caliper width = 0.2)",
                    "NNM",
                    "Covariate Adjustment",
                   "Crude")

# Convert Method column to factor with specified order
c$Method <- factor(c$Method, levels = desired_order)

my_plot <- ggplot(c,  aes(y = Method, x = Hazard_Ratio)) + geom_point(size = 2, color = "blue") + # Plot HR 
  geom_errorbarh(aes(xmin = Lower_CI, xmax = Upper_CI), height = 0.2) + # CI lines 
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.8) + # Reference line at HR = 1 
  labs(title = "Hazard Ratios for White vs. Asian ", x = "Hazard Ratio", y = "Method") + coord_cartesian(xlim = c(0.6,1.1)) + theme_minimal()

saveRDS(my_plot, "Asian_HR.rds")
```

```{r}
library(ggplot2)
library(gridExtra)
library(grid)
library(knitr)
library(kableExtra)

# Load and format data
df <- readRDS("Asian_table.rds") %>%
mutate(
Method = factor(Method, levels = rev(Method)),
CI = paste0("[", sprintf("%.2f", `95% CI Lower`), ", ", sprintf("%.2f", `95% CI Upper`), "]"),
HR_label = sprintf("%.2f", HR)
)

# Store y-axis levels
y_levels <- levels(df$Method)
n <- nrow(df)

# Table plot
table_plot <- ggplot(df, aes(y = Method)) +
geom_label(aes(x = 1, label = Method), hjust = 0, fill = NA, label.size = 0, size = 3.2, lineheight =2) +
geom_label(aes(x = 3.2, label = HR_label), hjust = 0, fill = NA, label.size = 0, size = 3.2, lineheight =2) +
geom_label(aes(x = 5.2, label = CI), hjust = 0, fill = NA, label.size = 0, size = 3.2, lineheight =2) +

# Column headers
annotate("text", x = 1, y = n + 1.8, label = "Method", fontface = "bold", hjust = 0, size = 3.5) +
annotate("text", x = 3.2, y = n + 1.8, label = "HR", fontface = "bold", hjust = 0, size = 3.5) +
annotate("text", x = 5.2, y = n + 1.8, label = "95% CI", fontface = "bold", hjust = 0, size = 3.5) +

scale_y_discrete(limits = y_levels) +
coord_cartesian(ylim = c(0.5, n + 2.2)) +
xlim(0.5, 6.5) +
theme_void()

# Forest plot
forest_plot <- ggplot(df, aes(x = HR, y = Method)) +
geom_errorbarh(aes(xmin = `95% CI Lower`, xmax = `95% CI Upper`), height = 0.1, color = "gray40") +
geom_point(shape = 15, size = 3, color = "blue") +
geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
scale_x_continuous(limits = c(0.8, 1.2)) +
scale_y_discrete(limits = y_levels) +
labs(x = "Hazard Ratio (HR) with 95% CI", y = NULL) +
theme_minimal(base_size = 11) +
theme(
axis.text.y = element_blank(),
axis.ticks.y = element_blank(),
panel.grid.major.y = element_blank()
)

# Combine plots
final_plot <- (table_plot + forest_plot) +
plot_layout(widths = c(3.5, 2)) +
plot_annotation(
title = "Hazard Ratios for White vs. Hispanic/Latino",
theme = theme(
plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
)
)

# Display
final_plot
```

```{r}
library(forestplot)
library(dplyr)
library(grid)
# Create your table data (assuming b is your original data.frame)
table_data <- data.frame(
Method = as.character(b$Method),
HR = sprintf("%.2f", b$HR),
CI = sprintf("[%.2f, %.2f]", b$`95% CI Lower`, b$`95% CI Upper`)
)

# Combine HR and CI into one column
table_data$HR_CI <- paste0(table_data$HR, " ", table_data$CI)

# Create label text matrix and add header
label_text <- rbind(
c("Method", "HR [95% CI]"),
cbind(table_data$Method, table_data$HR_CI)
)

# Plot
forestplot(
labeltext = label_text,
mean = c(NA, b$HR),
lower = c(NA, b$`95% CI Lower`),
upper = c(NA, b$`95% CI Upper`),
zero = 1,
boxsize = 0.2,
graph.pos = 3,
line.margin = 0.3,
lwd.zero = 2,
col = fpColors(box = "royalblue", line = "darkblue", zero = "black"),
xlab = "Hazard Ratio (HR) with 95% CI",
txt_gp = fpTxtGp(
label = gpar(fontsize = 11),
ticks = gpar(fontsize = 10),
xlab = gpar(fontface = "bold", fontsize = 12),
title = gpar(fontface = "bold", fontsize = 14)
),
title = "Hazard Ratios for White vs. Hispanic/Latino",
xticks = seq(0.6,1.2,by=0.1)
)

```

```{r}
# library(dplyr)
# library(forestplot)
# 
# # Step 1: Prepare table data
# table_data <- data.frame(
# Method = as.character(b$Method),
# HR = b$HR,
# logHR = b$loghr,
# SE = b$se
# )
# 
# # Step 2: Calculate 95% CI for HR scale
# table_data$Lower <- exp(table_data$logHR - 1.96 * table_data$SE)
# table_data$Upper <- exp(table_data$logHR + 1.96 * table_data$SE)
# 
# # Step 3: Create label text (left side) showing Method, logHR, logSE
# label_text <- cbind(
# table_data$Method,
# sprintf("%.2f", table_data$logHR),
# sprintf("%.2f", table_data$SE)
# )
# 
# # Step 4: Add header row
# label_text <- rbind(c("Method", "log(HR)", "SE"),
#                      label_text)
# is_summary <- c(TRUE,  rep(FALSE, nrow(table_data)))
# # Step 5: Bold only the header row
# label_styles <- fpTxtGp(
# label = list(
# gpar(fontface = c("plain", rep("plain", nrow(table_data)))),
# gpar(fontface = c("plain", rep("plain", nrow(table_data)))),
# gpar(fontface = c("plain", rep("plain", nrow(table_data))))
# ),
# ticks = gpar(fontsize = 10),
# xlab = gpar(fontsize = 11),
# title = gpar(fontface = "bold", fontsize = 14)
# )
# 
# # Step 6: Plot forestplot
# my_plot <- forestplot(
# labeltext = label_text,
# mean = c(NA,  table_data$HR),
# lower = c(NA,  table_data$Lower),
# upper = c(NA, table_data$Upper),
# zero = 1,
# boxsize = 0.2,
# line.margin = 0.3,
# is.summary = is_summary,
# lwd.zero = 2,
# graph.pos = 4, # 4th column for plot
# col = fpColors(box = "royalblue", line = "darkblue", zero = "black"),
# txt_gp = label_styles,
# xlab = "Hazard Ratio (HR) with 95% CI",
# title = "Race/Ethnicity Effect White vs. Hispanic/Latino",
# xticks = seq(0.6,1.1,by=0.1)
# )
# saveRDS(my_plot, "Asian_Forest.rds")


```

```{r}
library(dplyr)
library(forestplot)

# Step 1: Prepare table data
table_data <- data.frame(
Method = as.character(b$Method),
HR = sprintf("%.2f", b$HR),
CI = sprintf("[%.2f, %.2f]", b$`95% CI Lower`, b$`95% CI Upper`)
)

# Step 3: Create label text (left side) showing Method, logHR, logSE
label_text <- cbind(
table_data$Method,
table_data$HR,
 table_data$CI
)

# Step 4: Add header row
label_text <- rbind(c("Method", "HR", "95% CI"),
                     label_text)
is_summary <- c(TRUE,  rep(FALSE, nrow(table_data)))
# Step 5: Bold only the header row
label_styles <- fpTxtGp(
label = list(
gpar(fontface = c("plain", rep("plain", nrow(table_data))),fontsize = 11, fontfamily ="serif"),
gpar(fontface = c("plain", rep("plain", nrow(table_data))),fontsize = 11, fontfamily ="serif"),
gpar(fontface = c("plain", rep("plain", nrow(table_data))),fontsize = 11, fontfamily ="serif")
),
ticks = gpar(fontsize = 10, fontfamily ="serif"),
xlab = gpar(fontsize = 11, fontfamily ="serif"),
title = gpar(fontface = "bold", fontsize = 14)
)

# Step 6: Plot forestplot
my_plot <- forestplot(
labeltext = label_text,
mean = c(NA,  table_data$HR),
lower = c(NA, b$`95% CI Lower`),
upper = c(NA, b$`95% CI Upper`),
zero = 1,
boxsize = 0.2,
line.margin = 0.3,
is.summary = is_summary,
lwd.zero = 2,
graph.pos = 4, # 4th column for plot
col = fpColors(box = "royalblue", line = "darkblue", zero = "black"),
txt_gp = label_styles,
xlab = "Hazard Ratio (HR) with 95% CI",
xticks = seq(0.4,1.1,by=0.1)
)
saveRDS(my_plot, "Asian_Forest.rds")


```

##### Comparisons of Covariate Balance
```{r}
## Construct a data frame containing variable name and SMD from all methods
dataPlot <- data.frame(variable   = tabUnmatched$Variable,
                      Unmatched  = as.numeric(tabUnmatched$Mean_SMD),
                      Matched_optimal  = as.numeric(tabMatched_optimal_1$Mean_SMD),
                      Matched_caliper_0.1  = as.numeric(smd_avg_nn0.1_1$Mean_SMD),
                       Matched_nerest_neighbor = as.numeric(tabMatched_nn$Mean_SMD),
                       IPTW = as.numeric(smd_avg_iptw_1$Mean_SMD),
                                             IPTW_with_trimming = as.numeric(iptw_trimmed_1$Mean_SMD),
                      ps_stratification = as.numeric(weighted_smds$weighted_smd)
                       )

## Create long-format data for ggplot2
dataPlotMelt <- melt(data          = dataPlot,
                     id.vars       = c("variable"),
                     variable.name = "Method",
                     value.name    = "SMD")

library(forcats)

dataPlotMelt$Method <- fct_recode(dataPlotMelt$Method,
  "Original Data" = "Unmatched",
  "Optimal Matching" = "Matched_optimal",
  "NNM (caliper = 0.2)" = "Matched_caliper_0.1",
  "NNM" = "Matched_nerest_neighbor",
  "Stablized IPTW" = "IPTW",
  "IPTW (5% trimming)" = "IPTW_with_trimming",
  "PS Stratification" = "ps_stratification"
)

## Order variable names by magnitude of SMD
varNames <- as.character(dataPlot$variable)[order(dataPlot$Unmatched,decreasing = FALSE)]

## Order factor levels in the same order
dataPlotMelt$variable <- factor(dataPlotMelt$variable,
                                levels = varNames)

dataPlotMelt$variable2 <- fct_recode(dataPlotMelt$variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

## Plot using ggplot2
my_plot <- ggplot(data = dataPlotMelt,
mapping = aes(x = variable2, y = SMD, group = Method, color = Method)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 0.1, color = "red", linewidth = 0.3) +
coord_flip() +
theme_bw() +
theme(legend.key = element_blank(),
      ) +
scale_y_continuous(limits = c(0, 1)) +
labs(x = "Covariates", y = "SMD")

saveRDS(my_plot, file = "Asian_balance.rds")
```

```{r}
## Construct a data frame containing variable name and SMD from all methods
dataPlot <- data.frame(variable   = tabUnmatched$Variable,
                      Unmatched  = as.numeric(tabUnmatched$Mean_SMD)
                       )

## Create long-format data for ggplot2
dataPlotMelt <- melt(data          = dataPlot,
                     id.vars       = c("variable"),
                     variable.name = "Method",
                     value.name    = "SMD")

library(forcats)

dataPlotMelt$Method <- fct_recode(dataPlotMelt$Method,
  "Original Data" = "Unmatched"
)

## Order variable names by magnitude of SMD
varNames <- as.character(dataPlot$variable)[order(dataPlot$Unmatched,decreasing = FALSE)]

## Order factor levels in the same order
dataPlotMelt$variable <- factor(dataPlotMelt$variable,
                                levels = varNames)

dataPlotMelt$variable2 <- fct_recode(dataPlotMelt$variable,
  "SES Quintile" = "sesindex2015_2019",
  "Sex at Birth" = "gender",
  "Age at 1L Treatment" = "age_group",
  "Site of Care" = "practicetype",
  "Year of Diagnosis" = "year_diag_group",
  "Deletion 13q" = "Deletion.13q",
  "Anemia" = "hemoglobin",
  "ISS Stage" = "issstage",
  "ASCT" = "asct",
  "Cytogenetic Risk Group" = "risk_group",
  "Hypercalcemia" = "marker_calcium_result",
  "Ploidy" = "Ploidy",
  "ECOG Performance Status" = "combined_ecog",
  "Renal Insufficiency" = "marker_creatinie_result"
)

## Plot using ggplot2
my_plot <- ggplot(data = dataPlotMelt,
mapping = aes(x = variable2, y = SMD, group = Method, color = Method)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 0.1, color = "red", linewidth = 0.1) +
coord_flip() +
theme_bw() +
theme(legend.key = element_blank(),
      ) +
scale_y_continuous(limits = c(0, 1)) +
labs(x = "Covariates", y = "SMD")

```

# Average observations loss
```{r}
# Add method names to each
nnm_tbl <- match_summary_by_group_nnm %>%
mutate(Method = "NNM") %>% mutate(treatment = as.numeric(treatment)) %>% 
  rename(
Original_N = Avg_Original_N,
Retained_N = Avg_Matched_N,
Excluded_N = Avg_Excluded_N,
Prop_Lost = Avg_Prop_Excluded
)

nnm_caliper_tbl <- match_summary_by_group_nnmcaliper %>%
mutate(Method = "NNM (caliper = 0.2)") %>% mutate(treatment = as.numeric(treatment)) %>% 
  rename(
Original_N = Avg_Original_N,
Retained_N = Avg_Matched_N,
Excluded_N = Avg_Excluded_N,
Prop_Lost = Avg_Prop_Excluded
)

optimal_tbl <- match_summary_by_group_optimal %>%
mutate(Method = "Optimal Matching") %>% mutate(treatment = as.numeric(treatment)) %>% 
  rename(
Original_N = Avg_Original_N,
Retained_N = Avg_Matched_N,
Excluded_N = Avg_Excluded_N,
Prop_Lost = Avg_Prop_Excluded
)


trimming_tbl <- trim_summary_by_group %>%
mutate(Method = "IPTW (5% Trimming)") %>% 
  rename(
Original_N = Avg_Original_N,
Retained_N = Avg_Retained_N,
Excluded_N = Avg_Trimmed_N,
Prop_Lost = Avg_Prop_Trimmed
) %>% mutate(treatment1 = as.numeric(treatment)) %>% mutate(treatment = ifelse(treatment == 1, 1,0))


combined_tbl <- bind_rows(nnm_tbl, nnm_caliper_tbl, optimal_tbl, trimming_tbl) %>%
mutate(`Race/Ethnicity` = ifelse(treatment == 0, "White", "Asian")) %>%
select(Method, `Race/Ethnicity`, Original_N, Retained_N, Excluded_N, Prop_Lost)

cleaned_tbl <- combined_tbl %>%
mutate(
`Race/Ethnicity` = stringr::str_to_title(`Race/Ethnicity`),
`Prop Lost` = scales::percent(Prop_Lost, accuracy = 0.1)
) %>%
select(
Method,
`Race/Ethnicity`,
`Original N` = Original_N,
`Retained N` = Retained_N,
`Excluded N` = Excluded_N,
`Prop Lost`
) %>% mutate(across(c(`Original N`, `Retained N`, `Excluded N`), round, 0))

library(kableExtra)

cleaned_tbl %>%
kable(caption = "Table: Sample Sizes and Proportion Lost by Method and Group") %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

# Step 1: Prepare wide-style table with compact info
compact_tbl <- combined_tbl %>%
mutate(
Race = ifelse(`Race/Ethnicity` == "White", "White", "Asian"),
Loss_Info = paste0(round(Retained_N,0), " (", scales::percent(1-Prop_Lost, accuracy = 0.1), ")")
) %>%
select(Method, Race, Loss_Info) %>%
tidyr::pivot_wider(names_from = Race, values_from = Loss_Info) %>%
relocate(Method, White,`Asian`)



original_row <- tibble(
Method = "Original Average N",
White = round(nnm_tbl$Original_N[nnm_tbl$treatment == 0],0),
`Asian` = round(nnm_tbl$Original_N[nnm_tbl$treatment == 1],0)
) 

original_row$White <- as.character(original_row$White)
original_row$`Asian`<- as.character(original_row$`Asian`)

display_tbl <- bind_rows(original_row, compact_tbl)

a <- display_tbl %>%
kable(caption = "") %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

saveRDS(a, "Asian_sample loss.rds")
```


# Complete data analysis comparison
```{r}
completed_data <- read.csv("Asian_complete_data_results.csv")

# Load libraries 
library(ggplot2) 
library(dplyr) 

# Assuming you have the imputed results 
# Combine datasets for comparison 
combined_data <- bind_rows(completed_data %>% mutate(`Analysis Type` = "Complete Case"), c %>% mutate(`Analysis Type` = "Imputed Data") ) 
# Define the desired order manually
desired_order <- c("PS as Covariate",
                   "PS Stratification (5 strata)",
                 #  "PS Stratification (5 strata) - ATT",
                   "Stablized IPTW",
                #   "IPTW (5% trimming)",
                    "Optimal Matching",
                   "NNM (caliper width = 0.2)",
                    "NNM",
                    "Covariate Adjustment",
                   "Crude")

# Convert Method column to factor with specified order
combined_data$Method <- factor(combined_data$Method, levels = desired_order)

# Create a forest plot 
p <- ggplot(combined_data, aes(x = Method, y = Hazard_Ratio, color = `Analysis Type`)) + geom_point(position = position_dodge(width = 0.5), size = 3) + geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), position = position_dodge(width = 0.5), width = 0.3) + coord_flip() + labs( x = "Method", y = "HR with 95% CI") + scale_y_continuous(limits = c(0.4, 1.2)) + theme_minimal()

saveRDS(p, "Asian_comparison")
```

# SMD Prop

```{r, warning=FALSE, echo=FALSE}
library(cowplot)
# Add title above by stacking label + image
p1 <- plot_grid(
ggdraw() + draw_label("A. NNM (caliper = 0.2)", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
ggdraw() + draw_image("Asian_NNM02_SMD_Prop.png"),
ncol = 1,
rel_heights = c(0.1, 1)
)

p2 <- plot_grid(
ggdraw() + draw_label("B. NNM", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
ggdraw() + draw_image("Asian_NNM_SMD_Prop.png"),
ncol = 1,
rel_heights = c(0.1, 1)
)

p3 <- plot_grid(
ggdraw() + draw_label("C. Optimal Matching", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
ggdraw() + draw_image("Asian_optimal_SMD_Prop.png"),
ncol = 1,
rel_heights = c(0.1, 1)
)

p4 <- plot_grid(
ggdraw() + draw_label("D. Stabilized IPTW", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
ggdraw() + draw_image("Asian_SIPTW_SMD_Prop.png"),
ncol = 1,
rel_heights = c(0.1, 1)
)

# p5 <- plot_grid(
# ggdraw() + draw_label("E. IPTW (5% Trimming)", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
# ggdraw() + draw_image("Asian_IPTWTRIM_SMD_Prop.png"),
# ncol = 1,
# rel_heights = c(0.1, 1)
# )

p6 <- plot_grid(
ggdraw() + draw_label("F. PS Stratification", fontface = 'bold', x = 0.5, hjust = 0.5, size = 12),
ggdraw() + draw_image("Asian_STRATA_SMD_Prop.png"),
ncol = 1,
rel_heights = c(0.1, 1)
)

# Arrange in a 2x3 grid
combined <- plot_grid(p1, p2, p3, p4,p6, ncol = 2)

# Save combined panel
ggsave("Asian_combined_panel.png", combined, width = 10, height = 10)


```
















